% !TEX root = ../thesis-example.tex
%
\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}

I would like to take this opportunity to thank all the people without whom this study would never have been possible. Although it is just my name on the cover, it has been a collaborative effort as many have in their own special way helped and for that I want to give them special congratulations.

First, my supervisors Prof. Nikola Kasabov, Prof. Denise Taylor and Dr. Grace Wang. Nik, you have created the invaluable space for me to do this research and develop myself as a researcher in the best possible way. I greatly appreciate the freedom you have given me to find my own path, and the guidance and support you offered when needed. I highly appreciate you continuously pushing me to make this piece of work better. Denise and Grace, thank you for being an incredibly valuable part of this journey of mine and providing the guidance when required. I truly hope that you will give me the opportunity to work even closer together with you in the future. A special thanks also goes out to my examiners, Prof. Chrisina Jayne and Prof. Giacomo Indiveri for taking time to read and examine my thesis. 

I would also want to extend my gratitude towards Mrs. Joyce D'mello, the administrative manager of KEDRI. Joyce, you are the lifeline of KEDRI and I, among many, am indebted to you for all that you have done to make this journey worth. Thank you for helping me keep it all together in difficult times.  

Next, some people of outstanding importance for my research. Carolyn, Nathan and Anne, thank you for accepting to work alongside me, and for having numerous valuable conversations, all of which have shaken parts of my grey matter and resulted in this work. Our discussions were always lively, and although generally longer than planned, have always ended too soon. I hope to continue our collaborations in the future. I would also like to extend my acknowledgement to all the KEDRI members, especially Vivienne, Israel, Reggio, Fahad, Muhaini, Kaushalya, Clarence, Elisa, Maryam, Zohreh, Akshay, and everyone else I have had an opportunity to interact with. I also wish all of you all the very best for your future.   

I would not be sitting here today writing an acknowledgement for my thesis without the selfless support and encouragement of my family. Shumi, my dear wife, thank you for supporting me day in and day out throughout the ups and downs of this journey. This work is more yours than mine. Also, I thank you for your innumerable iterations of reading and proofreading this text. I am sure your PhD journey will much more enjoyable and you will always have my support on yours. Maa and baba, no words can ever describe my gratefulness to you and all the contributions you have made, quietly, to make this happen. Thank you for being my parents. I hope this work makes you proud. And of course to my precious little dogs, Burfi and Golu, who have always kept me going with all the love and cuddles. This writing is incomplete without a mention of all my family and friends in New Zealand, India and around the world who have continuously influenced me directly or indirectly over the years.

And finally, to my late maa and baba. This is for you. This is a gift that you deserve. May god bless you and may you rest in peace together. You are always and forever will be in my thoughts and prayers. 
